---
title: "Introduction to ipcwQR package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ipcwQR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.width = 7,
  fig.height = 4,
  fig.align = "center",
  dpi = 150,
  fig.path = "vignettes/ipcwQR-"
)
```

## Introduction

`ipcwQR` is the R package to fit the linear quantile regressions when the data are partially interval-censored and possibly correlated within same cluster.
Let $T$ and $X$ be the event time of interest and its related $p$-vector of covariates, respectively.
Our main objective is to estimate 
the $p$-dimensional quantile coefficient vector ${\boldsymbol{\beta}}_0(\tau)$
for some $\tau \in[\tau_L,\tau_R]\subset (0, 1)$ 
in the following linear quantile regression model:
$$
T_i = {\bf x}_i^T {\boldsymbol{\beta}}_0(\tau) + e_i(\tau),\quad i=1, \ldots ,n, 
$$
where $e_i(\tau)$ is the random error 
whose $\tau$th quantile conditional on ${\bf x}_i$ equals 0. 
When the data are subject to partially interval-censoring, 
left and right endpoints of the censoring time, $L$ and $R$,
are observed instead of $T$ such that $T\in(L,R)$.
Note that double-censoring  can also  be viewed as 
a special case of partly interval-censoring, 
i.e., $T$ is left-censored if $L=0$ and right-censored if $R=\infty$. 



## Usages

Installation of ipcwQR package can be done by

```{r}
devtools::install_github(repo="YejiStat/ipcwQR")
```

or
```{r}
base::require("ipcwQR")
```


picrq() function has the following arguments:
```{r}
# picrq(L=V,R=U,delta=delta,x=x,tau=tau)
```

see the detailed description from help(picrq()).

We first simulate univariate partly interval-censored data with normal random error, which is a similar simulation setting of Kim et al. (2022+).

```{r}
library(ipcwQR)
set.seed(316)
n = 200
x1 = runif(n,-1,1)
x2 = rbinom(n,1,0.43)
x = cbind(x1,x2)
T = 2 + x1 + x2 + rnorm(n)
U = (1 - 0.25*x1)*runif(n, -6, 5)
V = U + (1 - 0.1*x2)*runif(n, 6, 20)
U = exp(dplyr::case_when(TRUE ~ T, T>V ~ V, T<U ~ -Inf))
V = exp(dplyr::case_when(TRUE ~ T, T>V ~ Inf, T<U ~ U))
delta = ifelse(U==V, 1, 0)
tau=0.3
picrq(L=V,R=U,delta=delta,x=x,tau=tau)
picrq(L=V,R=U,delta=delta,x=x,tau=tau,estimation = "dr")
```

We posit two estimating methods, Gehan and log-rank, which can be conducted by specifying type = "gehan" and type = "logrank", respectively.

Next, we give illustrative example of the method for the multivariate clustered data, by using phase 3 metastatic colorectal cancer clinical trial. This dataset is available data(mCRC) in PICBayes package (Pan, 2021).

```{r}
library(ipcwQR)
library(tidyverse)
library(PICBayes)
data("mCRC")
d = with(data.frame(mCRC), data.frame(L = as.numeric(L),
                                      R = as.numeric(R),
                                      U = ifelse(y==0,R,L),
                                      V = ifelse(y==2,L,R),
                                      # Cluster weighted data
                                      id=(rep(c(table(SITE)),c(table(SITE)))),
                                      # Treatment arm: 0 = FOLFIRI alone, 1 = Panitumumab + FOLFIRI.
                                      x1= case_when(TRT_C == 0 ~ 0, #Pan et al data
                                                    TRT_C == 1 ~ 1),
                                      # Tumor KRAS mutation status: 0 = wild-type, 1 = mutant.
                                      x2= case_when(KRAS_C == 0 ~ 1,
                                                    KRAS_C == 1 ~ 0),
                                      site = as.numeric(SITE),
                                      y = as.numeric(y),
                                      delta = case_when(IC == 0 ~ 1,
                                                        IC == 1 ~ 0)
));
L=d$U;R=d$V; delta=d$delta
x = cbind(d$x1,d$x2); tau=0.3
picrq(d$U,d$V,d$delta,x=x,tau=tau)
picrq(d$U,d$V,d$delta,x=x,tau=tau, estimation = "dr")
picrq(d$U,d$V,d$delta,x=x,tau=tau,wttype = "nonparam",h=0.9)
picrq(d$U,d$V,d$delta,x=x,tau=tau,wttype = "nonparam",id=d$id,h=0.9)
```


## References

* Chiou, S. H., Kang, S. and Yan, J. (2015). 
Rank-based estimating equations with general weight for accelerated failure time models: an induced smoothing approach.
Statistics in Medicine 34(9): 1495–-1510.

* Pan, C. (2021). 
PICBayes: Bayesian Models for Partly Interval-Censored Data. R package. 
https://CRAN.R-project.org/package=PICBayes.

* Kim, Y., Choi, T., Park, S., Choi, S. and Bandyopadhyay, D. (2022+). 
Inverse weighted quantile regression with partially interval-censored data.
*Submitted to SMMR*.