---
title: "Introduction to ipcwQR package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ipcwQR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.width = 7,
  fig.height = 4,
  fig.align = "center",
  dpi = 150,
  fig.path = "vignettes/ipcwQR-"
)
```


## Introduction

`ipcwQR` is the R package to fit a new inverse-probability censoring weighted (IPCW) estimating procedure for censored quantile regression when the data are partially interval-censored that include doubly-censored (DC) data and partly interval-censored (PIC) and possibly correlated within the same cluster.
Let $T$ and $X$ be the event time of interest and its related $p$-vector of covariates, respectively.
Our main objective is to estimate 
the $p$-dimensional quantile coefficient vector ${\boldsymbol{\beta}}_0(\tau)$
for some $\tau \in[\tau_L,\tau_R]\subset (0, 1)$ 
in the following linear quantile regression model:
$$
T_i = {\bf x}_i^T {\boldsymbol{\beta}}_0(\tau) + e_i(\tau),\quad i=1, \ldots ,n, 
$$
where $e_i(\tau)$ is the random error 
whose $\tau$th quantile conditional on ${\bf x}_i$ equals 0. 
When the data are subject to partially interval-censoring, 
left and right endpoints of the censoring time, $L$ and $R$,
are observed instead of $T$ such that $T\in(L,R)$.
Note that double-censoring  can also  be viewed as 
a special case of partly interval-censoring, 
i.e., $T$ is left-censored if $L=0$ and right-censored if $R=\infty$. 



## Usages

Installation of ipcwQR package can be done by


```{r results='hide', message=FALSE, warning=FALSE}
devtools::install_github(repo="YejiStat/ipcwQR")
```

```{r results='hide', message=FALSE, warning=FALSE}
base::require("ipcwQR")
```


picrq() function has the following arguments:
```{r eval=FALSE, message=FALSE, warning=FALSE}
picrq(L=L,R=R,delta=delta,x=x,tau=tau)
```

drq() function has the following arguments:
```{r eval=FALSE, message=FALSE, warning=FALSE}
dcrq(L=L,R=R,T=T,delta=delta,x=x,tau=tau)
```

see the detailed description from help(picrq()) and help(dcrq()).

We first simulate univariate partly interval-censored data with normal random error, which is similar to PIC data simulation setting of Kim et al. (2022+).

```{r message=FALSE, warning=FALSE}
library(ipcwQR)
set.seed(111)
n = 200
x1 = runif(n,-1,1)
x2 = rbinom(n,1,0.43)
x = cbind(x1,x2)
T = 2 + x1 + x2 + rnorm(n)
U = (1 - 0.25*x1)*runif(n, -6, 5)
V = U + (1 - 0.1*x2)*runif(n, 6, 20)
U = exp(dplyr::case_when(TRUE ~ T, T>V ~ V, T<U ~ -Inf))
V = exp(dplyr::case_when(TRUE ~ T, T>V ~ Inf, T<U ~ U))
delta = ifelse(U==V, 1, 0)
tau=0.3
ipcwQR::picrq(L=V,R=U,delta=delta,x=x,tau=tau)
ipcwQR::picrq(L=V,R=U,delta=delta,x=x,tau=tau,estimation = "dr")
```


We posit two estimating methods, ipcw estimating method and doubly robust ipcw estimating method, which can be conducted by specifying estimation = NULL and estimation = "dr", respectively.

Next, we simulate univariate doubly-censored data with normal random error, which is one of DC data simulation setting of Kim et al. (2022+).

```{r message=FALSE, warning=FALSE}
n=200; tau=0.3
x1=runif(n,-1.2,1.7); x2=rbinom(n,1,0.6)
T = 1.7+x1+x2+rnorm(n)*(1-0.1*x2)
L=runif(n,-2.8,2); R=L+runif(n,4.2,8)
x=cbind(x1,x2)
delta=case_when(
    T<L ~ 3,
    T>R ~ 2,
    TRUE ~ 1 #observed
  )
ipcwQR::dcrq(L,R,T,delta,x,tau)
ipcwQR::dcrq(L,R,T,delta,x,tau,estimation = "dr")
ipcwQR::dcrq(L,R,T,delta,x,tau,wttype = "nonparam", hlimit = 0.9)
```



Finally, we simulate multivariate doubly-censored data with normal random error, which is one of MVDC data simulation setting of Kim et al. (2022+).

```{r message=FALSE, warning=FALSE}
cluster=60
v=rlnorm(cluster,mean=0,sd=0.5)
k = sample(0:9*10,cluster,replace = TRUE)
m = ifelse(quantile(v, k/100) <= v & quantile(v, (k+10)/100) <= v,
             k/10 + 2, 2)
id = as.numeric(unlist(apply(cbind(1:cluster,m),1,function(a) rep(a[1],a[2]))))
vi = as.numeric(unlist(apply(cbind(v,m),1,function(a) rep(a[1],a[2]))))
n=sum(m)
x1=runif(n,-1.4,0.9); x2=rbinom(n,1,0.56)
T = 2.3+x1+x2+rnorm(n)*x2*0.1+vi*rnorm(n)*0.85
L=runif(n,-2.8,2); R=L+runif(n,4.2,9.7)
x=cbind(x1,x2)
delta=case_when(
    T<L ~ 3,
    T>R ~ 2,
    TRUE ~ 1 #observed
  )
dcrq(L,R,T,delta,x,tau,id=NULL)
dcrq(L,R,T,delta,x,tau,id=id)
```



Next, we apply our proposed method to the multivariate clustered data using a phase 3 metastatic colorectal cancer clinical trial. This dataset is accessible for data(mCRC) in the PICBayes R package (Pan, 2021).

```{r message=FALSE, warning=FALSE}
library(PICBayes)

data("mCRC")
d = with(data.frame(mCRC), data.frame(L = as.numeric(L),
                                      R = as.numeric(R),
                                      U = ifelse(y==0,R,L),
                                      V = ifelse(y==2,L,R),
                                      # Cluster weighted data
                                      # Treatment arm: 0 = FOLFIRI alone, 1 = Panitumumab + FOLFIRI.
                                      x1= case_when(TRT_C == 0 ~ 0, #Pan et al data
                                                    TRT_C == 1 ~ 1),
                                      # Tumor KRAS mutation status: 0 = wild-type, 1 = mutant.
                                      x2= case_when(KRAS_C == 0 ~ 1,
                                                    KRAS_C == 1 ~ 0),
                                      id = as.numeric(SITE),
                                      y = as.numeric(y),
                                      delta = case_when(IC == 0 ~ 1,
                                                        IC == 1 ~ 0)
));
L=(log(d$U));R=log(d$V); delta=d$delta
x = cbind(d$x1,d$x2); id=d$id;  tau=0.1;

ipcwQR::picrq(L,R,delta,x=x,tau=tau)
ipcwQR::picrq(L,R,delta,x=x,tau=tau, estimation = "dr")
ipcwQR::picrq(L,R,delta,x=x,tau=tau,id=id,hlimit=0.9,k=2)
ipcwQR::picrq(L,R,delta,x=x,tau=tau,id=NULL,hlimit=0.9)
```


## References

* Beran, R. (1981). Nonparametric Regression with Randomly Censored Survival Data. Technical Report, Univ.California, Berkeley.

* Chiou, S. H., Kang, S. and Yan, J. (2015). 
Rank-based estimating equations with general weight for accelerated failure time models: an induced smoothing approach.
Statistics in Medicine 34(9): 1495–-1510.

* Pan, C. (2021). 
PICBayes: Bayesian Models for Partly Interval-Censored Data. R package. 
https://CRAN.R-project.org/package=PICBayes.

* Kim, Y., Choi, T., Park, S., Choi, S. and Bandyopadhyay, D. (2022+). 
Inverse weighted quantile regression with partially interval-censored data.
*Submitted to SMMR*.